---
title: "Effects of Cationic Polyacrylamide and Cationic Starch on Aquatic Life"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dslabs)
library(tidyverse)
```
<br>

## Abstract

Geotextile tubes with polyacrylamide flocculants are widely used in dewatering applications. Due to variations in solid concentrations during dredging, excess flocculant is sometimes released into the environment, where it might have toxic effects. This study determined optimum doses for a cationic polyacrylamide (CPAM) and a natural-based polymer alternative, cationic starch (C. Starch). Slurry samples were treated with optimum and 50% overdoses of each compound, and residual polymer concentrations were measured. Overdosed C. Starch resulted in low residuals (<2 ppm), but overdosed CPAM resulted in 17.4 ppm residual polymer. The relative toxicity of CPAM and C. Starch was also tested using zebrafish embryos. 100% of embryos that had their chorion removed and 71.8% of embryos that retained their chorions, were dead or dying after 7 days of exposure to CPAM. In contrast, there was no statistically significant difference in the numbers of embryos that were dead or dying, when exposed to C. Starch, compared to controls. These data strongly suggest that C. Starch should be considered as a replacement to CPAM in dewatering applications.

<br>
**Keywords:** Geosynthetics, geotextile tube, dewatering, cationic starch, cationic polyacrylamide, toxicity, zebrafish, lethality

<br>

### Authors

* **Katie L. Duggan:** Undergraduate Researcher, Department of Civil & Environmental Engineering, Syracuse University, Syracuse NY
* **Martina Morris:** Undergraduate Researcher, Department of Biology, Syracuse University, Syracuse NY
* **Shobha K. Bhatia:** Laura J. and L Douglas Meredith Professor for Teaching Excellence, Department of Civil & Environmental Engineering, Syracuse University, Syracuse NY
* **Mahmoud M. Khachan:** Graduate Research Assistant, Department of Civil & Environmental Engineering, Syracuse University, Syracuse NY
* **Katharine E. Lewis:** Professor of Biology, Department of Biology, Syracuse University, Syracuse NY

<br>

### Source:

Duggan, K. L., Morris, M., Bhatia, S. K., Khachan, M. M., & Lewis, K. E. (2019). Effects of Cationic Polyacrylamide and Cationic Starch on Aquatic Life. Journal of hazardous, toxic and radioactive waste, 23(4), 10.1061/(ASCE)HZ.2153-5515.0000467. https://doi.org/10.1061/(ASCE)HZ.2153-5515.0000467